Sequential Decision-Making under Uncertainty with Big Data

### 2023-03-30 paper/main.tex
merged

%\title{Investigatin Uncertainty Quantification in Model-based Reinforcement Learning}
\title{Model-based Reinforcement Learning with Fast Posterior Updates}
\author{%
  Aidan ~Scannell \\
  Aalto University\\
  %Finnish Center for Artificial Intelligence \\
  \texttt{aidan.scannell@aalto.fi}
  \And
  Paul ~Chang \\
  Aalto University\\
  \texttt{paul.chang@aalto.fi}
  \And
  Ella ~Tamir \\
  Aalto University\\
  \texttt{ella.tamir@aalto.fi}
  \And
  Arno ~Solin \\
  Aalto University\\
  \texttt{arno.solin@aalto.fi}
  \And
  Joni ~Pajarinen \\
  Aalto University\\
  \texttt{joni.pajarinen@aalto.fi}
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

@@

% Domains
\newcommand{\stateDomain}{\ensuremath{\mathcal{S}}}
\newcommand{\actionDomain}{\ensuremath{\mathcal{A}}}
\newcommand{\policyDomain}{\ensuremath{\Pi}}

@@

\begin{abstract}
  % Reinforcement learning (RL) agents typically do not utilise the information previously gained during an episode.
  % as they perform learning offline, i.e. after an episode.
  % Reinforcement learning (RL) agents typically perform learning after an episode, ignoring the potential information gain whilst collecting data.
  % Reinforcement learning (RL) agents typically perform learning offline and do not utilise the information gained during an episode.
  % We present a Bayesian model-based RL algorithm which incorporates this information by updating a dynamic model's posterior during an episode.
  % Our method leverages the fast posterior updates available in the dual parameterisation of sparse variational Gaussian processes.
  % Importantly, our algorithm scales to 1) environment's with high dimensional state-spaces as it is built upon the connection between Bayesian neural networks (BNNs) and
  % Gaussian processes (GPs) and 2) long episodes, as the complexity of our updates is independent of episode length.
  % We demonstrate that our fast posterior updates improve sample efficiency in environments with action penalties, a notoriously difficult
  % challenge for model-based RL algorithms.
  Reinforcement learning (RL) agents typically perform learning offline and do not utilise the information gained during an episode.
  We present a Bayesian model-based RL algorithm which incorporates this information by updating a dynamic model's posterior during an episode.
  Typically training a neural network dynamic model is not feasible and nonparameteric methods such as Gaussian processes scale poorly with data.
  Importantly, our algorithm scales to 1) environment's with high dimensional state-spaces, 2) large data sets and 3) long episodes,
  as the complexity of our updates is independent of episode length.
  We demonstrate that our fast posterior updates improve sample efficiency in environments with action penalties, a notoriously difficult
  challenge for model-based RL algorithms.
\end{abstract}

@@

Why model-based RL:
\begin{itemize}
  \item Better sample efficiency than model-free RL
\end{itemize}

Why Bayesian model-based RL:
\begin{itemize}
  \item Can be used to combat model bias
  \item Better exploration in environment's that are hard to explore (sparse reward/control penalties)
  \item Bayesian framwork can naturally handle stochastic environments
\end{itemize}

Issues with uncertainty in model-based RL:
\begin{itemize}
  \item Over exploration when using optimism in the face of uncertainty
  \begin{itemize}
    \item Agent drives exploration into region of high epistemic uncertainty because it thinks it could have high return.
    \item In reality, the dynamic model is innacurate and the agent spends the majority of an episode exploiting this innacuracy.
    \item What if the agent could update it's model during an episode and realise it's pointless to explore in this region
    \item We show that this can reduce model exploitation and in turn, improve sample efficiency.
  \end{itemize}
\end{itemize}

\section{Problem Statement and Background} \label{sec:problem-statement}
We consider environments with states \(\state \in \stateDomain \), actions \(\action \in \actionDomain\) and transition dynamics \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \) subject to
iid noise \(\noise_{t}\), given by,
\begin{align}
\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}.
\end{align}
\begin{assumption}
  (system properties) The environment's dynamics are $L_{\transitionFn}\text{-Lipschitz}$ continuous and the transition noise $\noise_{t}$ is $\sigma\text{-sub-Gaussian}$ for all $t \geq 0$.
\end{assumption}

>>>>

%\title{Investigatin Uncertainty Quantification in Model-based Reinforcement Learning}
% \title{Model-based Reinforcement Learning with Fast Posterior Updates}
\title{Sequential Decision-Making under Uncertainty with Big Data}
\author{%
  Aidan ~Scannell \\
  Aalto University\\
  %Finnish Center for Artificial Intelligence \\
  \texttt{aidan.scannell@aalto.fi}
  \And
  Paul ~Chang \\
  Aalto University\\
  \texttt{paul.chang@aalto.fi}
  \And
  Ella ~Tamir \\
  Aalto University\\
  \texttt{ella.tamir@aalto.fi}
  \And
  Arno ~Solin \\
  Aalto University\\
  \texttt{arno.solin@aalto.fi}
  \And
  Joni ~Pajarinen \\
  Aalto University\\
  \texttt{joni.pajarinen@aalto.fi}
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

@@

% Domains
\newcommand{\stateDomain}{\ensuremath{\mathcal{S}}}
\newcommand{\actionDomain}{\ensuremath{\mathcal{A}}}
\newcommand{\inputDomain}{\ensuremath{\mathbb{R}^{D}}}
\newcommand{\outputDomain}{\ensuremath{\mathbb{R}^{C}}}
\newcommand{\policyDomain}{\ensuremath{\Pi}}

@@

\begin{abstract}
  % Reinforcement learning (RL) agents typically do not utilise the information previously gained during an episode.
  % as they perform learning offline, i.e. after an episode.
  % Reinforcement learning (RL) agents typically perform learning after an episode, ignoring the potential information gain whilst collecting data.
  % Reinforcement learning (RL) agents typically perform learning offline and do not utilise the information gained during an episode.
  % We present a Bayesian model-based RL algorithm which incorporates this information by updating a dynamic model's posterior during an episode.
  % Our method leverages the fast posterior updates available in the dual parameterisation of sparse variational Gaussian processes.
  % Importantly, our algorithm scales to 1) environment's with high dimensional state-spaces as it is built upon the connection between Bayesian neural networks (BNNs) and
  % Gaussian processes (GPs) and 2) long episodes, as the complexity of our updates is independent of episode length.
  % We demonstrate that our fast posterior updates improve sample efficiency in environments with action penalties, a notoriously difficult
  % challenge for model-based RL algorithms.
  %
  % Reinforcement learning (RL) agents typically perform learning offline and do not utilise the information gained during an episode.
  % We present a Bayesian model-based RL algorithm which incorporates this information by updating a dynamic model's posterior during an episode.
  % Typically training a neural network dynamic model is not feasible and nonparameteric methods such as Gaussian processes scale poorly with data.
  % Importantly, our algorithm scales to 1) environment's with high dimensional state-spaces, 2) large data sets and 3) long episodes,
  % as the complexity of our updates is independent of episode length.
  % We demonstrate that our fast posterior updates improve sample efficiency in environments with action penalties, a notoriously difficult
  % challenge for model-based RL algorithms.
  %
  Deep neural networks

  Obtain SVGP posterior on trained deep neural network (NN) without optimising any variational parameters.

  Dual formulation enables fast updates for 1) model-based RL 2) pathwise, function-space conditioning.
  % Reinforcement learning (RL) agents typically perform learning offline and do not utilise the information gained during an episode.
  % We present a Bayesian model-based RL algorithm which incorporates this information by updating a dynamic model's posterior during an episode.
  % Typically training a neural network dynamic model is not feasible and nonparameteric methods such as Gaussian processes scale poorly with data.
  % Importantly, our algorithm scales to 1) environment's with high dimensional state-spaces, 2) large data sets and 3) long episodes,
  % as the complexity of our updates is independent of episode length.
  % We demonstrate that our fast posterior updates improve sample efficiency in environments with action penalties, a notoriously difficult
  % challenge for model-based RL algorithms.
\end{abstract}

@@

\begin{itemize}
  \item DNNs scale to high-dimensional and large data sets but 1) no UQ and 2) cannot be updated quickly
  \begin{itemize}
    \item In many domains we want to make decisions under a learned model's uncertainty, e.g. in healthcare, robotics, autonomous vehicles. (safey, sample efficiency (cost/time))
    \item In continual learning (e.g. online learning/RL) we need to update model quicker than retraining NN with SGD.
    \item Poorly calibrated model uncertainty
  \end{itemize}
\end{itemize}

\begin{itemize}
  \item BNNs give UQ over high-dimensional parameter space,
  \begin{itemize}
    \item Exact inference intractable
    \item Sampling in parameter space vs Sampling in function space
  \end{itemize}
\end{itemize}

\begin{itemize}
  \item BNN to dual SVGP
  \begin{itemize}
    \item Keep in spirit of Laplace approx and fit SVGP with no more optimisation
    \item Now we can update posterior without retraining, which is fast (good for model-based RL and TS BO?)
    \item Now we can sample in function space with pathwise conditioning (because we have SVGP).
    \begin{itemize}
      \item Our pathwise conditioning will be better as we can use dual update
    \end{itemize}
  \end{itemize}
\end{itemize}

\section{Background} \label{sec:background}
% \section{dual SVGP}
% In this section we extend these fast updates to environment's with high dimensional state spaces and large data sets.
% % Our method draws on the connection between BNNs and GPs and formulates a function space SVGP posterior given
% Our method uses a BNN dynamic model and draws on the connection between BNNs and GPs
% \parencite{khanApproximate2019} to formulate a function space SVGP posterior,
% where we can apply the fast updates from \cref{eq-dual-update-svgp}.
% At a high-level, we first use Laplace's approximation to obtain a weight space posterior for our BNN.
% We then linearise our BNN around the optimal parameters and interpret it as a GP.
% Finally, we formulate a lower rank approximation of this GP posterior (i.e. a SVGP posterior) using inducing variables.
% % formulates a function space SVGP posterior by drawing on the connection between BNNs and GPs.

In this section, we recap how to fit a GP posterior to a trained deep neural network (DNN).
We start by obtaining a posterior over the weights of a trained neural network via the Laplace approximation.
We then show how we can obtain a function-space posterior, (i.e. a GP posterior)
by linearising the neural network around the weights Maximum a Posteriori (MAP) estimate.
% We then show that by linearising the neural network we can obtain a function-space posterior, i.e. a GP posterior.

% We paraterise our single-step dynamic model $\transitionFnWithParams : \inputDomain \rightarrow \outputDomain$
% as an $L\text{-layer}$ NN with weights $\transitionParams$.
Given a data set $\dataset = \{\mathbf{x}_{n} \in \inputDomain, \mathbf{y}_{n} \in \outputDomain\}_{n=0}^{N}$,
the goal of (supervised) deep learning, is to train the weights $\transitionParams \in \R^{P}$ of an $L\text{-layer}$ NN
$f : \inputDomain \rightarrow \outputDomain$ to minimize the (regularized) empirical risk,
\begin{align}
  \transitionParams^{*} = \arg \min_{\transitionParams \in \R^{D}} \mathcal{L}(\dataset;\transitionParams) = \arg \min_{\transitionParams \in \R^{D}} \left(
  R(\transitionParams) + \sum_{n=0}^{N-1} l(f(\mathbf{x}_{n} ; \transitionParams), \mathbf{y}_{n}) \right).
\end{align}
In practice, it is common to use the mean squared error (MSE) loss
$l(f(\mathbf{x}_{n} ; \transitionParams),\mathbf{y}_{n}) = \|f(\mathbf{x}_{n} ; \transitionParams) - \mathbf{y}_{n} \|^{2}_{2}$
and to use the weight decay regularizer $R(\transitionParams)=\frac{1}{2}\gamma^{-2}\|\transitionParams\|^{2}_{2}$.
However, the practioner is free to choose any loss and regularizer. \todo{I think?}

\textbf{Laplace approximation}
The Laplace approximation \todo{add citation of original paper} constructs a Gaussian approximation of the weight-space posterior $p(\transitionParams \mid \dataset)$
by using a second-order Taylor expansion of $\mathcal{L}$ around $\transitionParams^{*}$.
The posterior approximation is given by,
\begin{align} \label{eq-laplace-approx-weight-space}
  p(\transitionParams \mid \dataset) \approx \mathcal{N} \left( \transitionParams \mid \transitionParams^{*} , \bm\Sigma_{\transitionParams^{*}} \right)
  \quad \text{with} \quad \bm\Sigma_{\transitionParams} =
 - \nabla_{\transitionParams \transitionParams}^{2} \mathcal{L} ( \dataset ; \transitionParams)|_{\transitionParams=\transitionParams^{*}}
\end{align}
That is, it sets the posterior precision to the Hessian of the loss at the optimal parameters $\transitionParams^{*}$.
It is worth noting that computing the Hessian of the loss can be computationally intractable for large networks.
The prior terms are usually trivial, so we will focus on the likelihood here.
The Jacobian and Hessian of the log likelihood can be expressed per data point,
\begin{align} \label{eq-jac}
  \nabla_{\transitionParams} \log p(\mathbf{y} \mid \mathbf{f}(\mathbf{x}; \transitionParams)) &= \mathbf{J}(\mathbf{x})^{T} \mathbf{r}(\mathbf{y} ; \mathbf{f}) \\
  \nabla^{2}_{\transitionParams\transitionParams} \log p(\mathbf{y} \mid \mathbf{f}(\mathbf{x}; \transitionParams)) &= \mathbf{H}(\mathbf{x})^{T}
  \mathbf{r}(\mathbf{y};\mathbf{f}) - \mathbf{J}(\mathbf{x})^{T} \bm\Lambda(\mathbf{y};\mathbf{f}) \mathbf{J}(\mathbf{x}),
\label{eq-hess}
\end{align}
through the Jacobian $\mathbf{J} \in \R^{C \times P}$ and Hessian $\mathbf{H} \in \R^{C \times P \times P}$ of the feature extractor $\mathbf{f}(\mathbf{x}; \bm\transitionParams)$,
\begin{align} \label{eq-jac}
[\mathbf{J}(\cdot)]_{ci} = \odv{f_{c}(\cdot ; \transitionParams)}{\transitionParams_{i}}_{\transitionParams=\transitionParams^{*}} \qquad
[\mathbf{H}(\cdot)]_{cij} = \odv{f_{c}(\cdot ; \transitionParams)}{\transitionParams_{i}\transitionParams_{j}}_{\transitionParams=\transitionParams^{*}}
\end{align}
where
$\mathbf{r}(\mathbf{y}; \mathbf{f}) = \nabla_{\mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$ can be interpreted as a residual and
$\bm\Lambda(\mathbf{y} ; \mathbf{f}) &= \nabla^{2}_{\mathbf{f} \mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$
as per-input noise.
Many approximations exist and can be used alongsied our method.
We refer the reader to \cite{daxbergerLaplace2021} for more details.
See \cref{sec-laplace-approximation} for further details on the Laplace approximation.

% \begin{align}
% \Sigma_{\transitionParams_{i}}^{-1} = - \nabla_{\transitionParams \transitionParams}^{2} \log p(\transitionParams \mid \dataset_{0:i})|_{\transitionParams=\transitionParams^{*}_{i}}
%   + \frac{1}{\sigma^{2}_{0}} \mathbf{I}
% \end{align}

\textbf{Weight space to function space}
% We can represent the log joint density in the function space,
% \begin{align} \label{eq-laplace-approx-function-space}
%   \sum_{n=1}^{N} \log p \left(y_{n} \mid f(\mathbf{x}_{n} ; \transitionParams_{0:i}) \right) + \log p(f(\mathbf{X}; \transitionParams_{0:i}))
% \end{align}
We can represent the weight space posterior from \cref{eq-laplace-approx-weight-space}, in
function space, by linearising the BNN and interpretting it as a GP.
The GP prior is then given by,
\begin{align} \label{eq-laplace-approx-function-space}
  % \transitionFn_{\transitionParams_{i}}(\cdot) \sim \mathcal{N} \left( \mu_{\transitionParams_{i}^{*}}(\cdot), K_{\transitionParams_{i}^{*}}(\cdot, \cdot') \right) \qquad
  f(\cdot ;\transitionParams) \sim \mathcal{N} \left( \mu(\cdot), k(\cdot, \cdot') \right) \qquad
  % \mu_{\transitionParams_{i}^{*}}(\cdot)
  \mu(\cdot)
  = f(\cdot ;\transitionParams^{*}) \qquad
  % K_{\transitionParams_{i}^{*}}(\cdot, \cdot')
  k(\cdot, \cdot')
  = \gamma^{2} \mathbf{J}(\cdot) \mathbf{J}(\cdot')^{T},
\end{align}
where the kernel is the neural tangent kernel (NTK) \parencite{immerImprovingPredictionsBayesian2021}.
\todo{what mean function shoudl this use???}
\todo{NTK times some factor??}
% with Jacobian given by $\mathbf{J}(\cdot) = \odv{\transitionFnWithParams(\cdot)}{\transitionParams}_{\transitionParams=\transitionParams_{i}^{*}}$.
% The kernel in \cref{eq-laplace-approx-function-space} is the neural tangent kernel (NTK)
% $k(\mathbf{x}, \mathbf{x}') = \sigma_{0}^{2} J_{\transitionParams^{*}}(\mathbf{x}) J_{\transitionParams^{*}}(\mathbf{x}')^{T}$
% \parencite{immerImprovingPredictionsBayesian2021}
% to formulate the GP posterior $p(f(\mathbf{X} ;\transitionParams_{0:i}) \mid \dataset_{0:i})$.
It is worth noting that the NTK linearises the network around the optimised parameters $\transitionParams_{i}^{*}$,
so the function space prior (and thus posterior) is only a locally linear approximation.
In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
It is also worth noting that the NTK may be highly non stationary and is dependent on the NN architecture, for example, the activation functions.
Given our GP prior, we can use the properties of multivariate normals to obtain our GP posterior by conditioning on the data.
Our GP posterior is given by,
\todo{add mean func to GP posterior if we're using one}
\begin{align} \label{eq-gp-predictive-posterior}
  p(f(\mathbf{x}_{*}) \mid \mathbf{y}) &= \mathcal{N}
  % \left( f(\mathbf{x}_{*}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{x}\mathbf{x}}^{-1} \mathbf{A}^{T} \right) \\
  \left( f(\mathbf{x}_{*}) \mid \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}} \mathbf{m}^{*},
  k_{**} - \mathbf{k}_{*\mathbf{x}} \left(\mathbf{K}_{\mathbf{x}\mathbf{x}} + \bm\Lambda(\mathbf{y} ; \mathbf{f})^{-1} \right)^{-1} \mathbf{k}_{*\mathbf{x}}^{T} \right)
\end{align}

\section{Dual SVGP} \label{sec:method}
The GP predictive posterior in \cref{eq-gp-predictive-posterior} is computationally expensive
as it requires inverting $N\times N$ matrix.
Sparse approximations using pseudo points ...
\begin{align} \label{eq-svgp-predictive-posterior}
  p(f(\mathbf{x}_{*}) \mid \mathbf{y}) &= \mathcal{N}
  % \left( f(\mathbf{x}_{*}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{x}\mathbf{x}}^{-1} \mathbf{A}^{T} \right) \\
  \left( f(\mathbf{x}_{*}) \mid \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}} \mathbf{m}^{*},
  k_{**} - \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}} \mathbf{k}_{*\mathbf{x}}^{T} \right) \\
  % &\approx \mathcal{N}
  % \left( f(\mathbf{x}_{*}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{z}\mathbf{z}}^{-1} \mathbf{A}^{T} + \mathbf{A} \mathbf{V}^{*} \mathbf{A}^{T} \right) \\
&\approx \mathcal{N} \left( f(\mathbf{x}_{*}) \mid \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{m}^{*},
  k_{**} - \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\mathbf{z}}^{T}
  % \mathbf{k}_{*\inducingInput} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\inducingInput}^{T}
  + \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}}  \mathbf{V}^{*}  \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\mathbf{z}}^{T}
  \right)
\coloneqq q_{\inducingVariable}(f(\mathbf{x}_{*}))
\end{align}
% where $\mathbf{A} = \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}}$

We follow \cite{csatoSparseOnlineGaussian2002} and express our GP posterior in the dual parameter space.
However, we consider the variational formulation introduced by \cite{adamDualParameterizationSparse2021}, where the optimal variational parameters are given by,
\begin{align} \label{eq-dual-params}
\mathbf{m}^{*} = \mathbf{V}^{*}\bm\alpha^{*} \quad \mathbf{V}^{*} = [\mathbf{K}_{\mathbf{z}\mathbf{z}}^{-1} + \bm\beta^{*}]
\end{align}
Importantly, we can calculate the natural parameters $(\bm\alpha, \bm\beta)$ in closed form,
\todo{paul needs to update \cref{eq-dual-svgp-params}}
\begin{align} \label{eq-dual-svgp-params}
\bm\alpha_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right] \\
\bm\beta_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right].
\end{align}
This simplifies SVGP inference \parencite{hensmanGaussian2013} as it no longer requires the variational parameters $(\mathbf{m}, \mathbf{V})$ to be optimised.

% Importantly, \cite{changFantasizingDualGPs2022} show that in the dual space,
% conditioning on new observations reduces to suming the dual variables from the previous time
% step with an update, given by,
% \todo{pick best way to show new data (using superscript new or just time indexing?)}
% \begin{align} \label{eq-dual-update-svgp}
% \bm\alpha_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right] \\
% \bm\beta_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right]
% \end{align}

% We follow \cite{csatoSparseOnlineGaussian2002} and express our GP posterior in the dual parameter space,
% % to write GP posterior $p(f(\mathbf{X} ;\transitionParams_{0:i}) \mid \dataset_{0:i})$ as,
% \begin{align} \label{eq-gp-post}
%   \mathbb{E}_{p(f(\mathbf{x}_{*}) \mid \mathbf{y}_{n})} \left[ f(\mathbf{x}_{*}) \right] &= \mu(\mathbf{x}_{*}) + \mathbf{k}_{\mathbf{x}*}^{T} \bm\alpha_{n} \\
%   \mathbb{V}_{p(f(\mathbf{x}_{*}) \mid \mathbf{y}_{n})} \left[ f(\mathbf{x}_{*}) \right] &= k_{**} - \mathbf{k}_{\mathbf{x}*}^{T} \left( \mathbf{K}_{\mathbf{x} \mathbf{x}} + \text{diag}(\bm\beta_{n})^{-1} \right)^{-1} \mathbf{k}_{\mathbf{x}*}
% \end{align}
% % \begin{align} \label{eq-laplace-approx-function-space}
% %   \mathbb{E}_{p(f_{*} \mid \mathbf{y}_{n})} \left[ f_{*} \right] &= \mathbf{k}_{\mathbf{x}*}^{T} \bm\alpha_{n} \\
% %   \mathbb{V}_{p(f_{*} \mid \mathbf{y}_{n})} \left[ f_{*} \right] &= k_{**} - \mathbf{k}_{\mathbf{x}*}^{T} \bm\alpha_{n} \left( \mathbf{K}_{\mathbf{x} \mathbf{x}} + \diag(\bm\beta_{n})^{-1} \right)^{-1} \mathbf{k}_{\mathbf{x}*}
% % \end{align}
% where the dual parameters $(\bm\alpha_n, \bm\beta_n)$ are vectors of,
% \begin{align} \label{eq-gp-dual-params}
%   \bm\alpha_{n} \coloneqq \left\{ \mathbb{E}_{p(f(\mathbf{x}_{i}) \mid \mathbf{y}_{n})} \left[ \nabla_{f_{i}} \log p(y_{i} \mid f(\mathbf{x}_{i})) \right] \right\}_{i=1}^{N} \qquad
%   \bm\beta_{n} \coloneqq \left\{ \mathbb{E}_{p(f(\mathbf{x}_{i}) \mid \mathbf{y}_{n})} \left[ \nabla^{2}_{f_{i}} \log p(y_{i} \mid f(\mathbf{x}_{i})) \right] \right\}_{i=1}^{N}
% \end{align}
% Importantly, this parameterisation

% \textbf{dual-SVGP}
% \begin{align} \label{eq-svgp-post}
%   \mathbb{E}_{q_{\mathbf{u}}(f(\mathbf{x}_{i}))} \left[ f(\mathbf{x}_{*}) \right] &= \mu(\mathbf{x}_{*}) + \mathbf{k}_{\mathbf{z}*}^{T} \bm\alpha_{\mathbf{u}} \\
%   \mathbb{V}_{q_{\mathbf{u}}(f(\mathbf{x}_{i}))} \left[ f(\mathbf{x}_{*}) \right] &= k_{**} - \mathbf{k}_{\mathbf{z}*}^{T} \left[ \mathbf{K}_{\mathbf{z}\mathbf{z}}^{-1} - \left( \mathbf{K}_{\mathbf{z} \mathbf{z}} + \bm\beta_{\mathbf{u}} \right)^{-1} \right] \mathbf{k}_{\mathbf{z}*}
% \end{align}
% % \begin{align} \label{eq-laplace-approx-function-space}
% %   \mathbb{E}_{p(f_{*} \mid \mathbf{y}_{n})} \left[ f_{*} \right] &= \mathbf{k}_{\mathbf{x}*}^{T} \bm\alpha_{n} \\
% %   \mathbb{V}_{p(f_{*} \mid \mathbf{y}_{n})} \left[ f_{*} \right] &= k_{**} - \mathbf{k}_{\mathbf{x}*}^{T} \bm\alpha_{n} \left( \mathbf{K}_{\mathbf{x} \mathbf{x}} + \diag(\bm\beta_{n})^{-1} \right)^{-1} \mathbf{k}_{\mathbf{x}*}
% % \end{align}
% where the dual parameters $(\bm\alpha_{\mathbf{u}}, \bm\beta_{\mathbf{u}})$ are vectors of,
% \begin{align} \label{eq-svgp-dual-params}
%   \bm\alpha_{\mathbf{u}} \coloneqq \mathbf{k}_{\mathbf{z}i} \mathbb{E}_{q_{\mathbf{u}}(f(\mathbf{x}_{i}))} \left[ \nabla_{f_{i}} \log p(y_{i} \mid f(\mathbf{x}_{i})) \right] \qquad
%   \bm\beta_{\mathbf{u}} \coloneqq \mathbf{k}_{\mathbf{z}i} \mathbb{E}_{q_{\mathbf{u}}(f(\mathbf{x}_{i}))} \left[ \nabla^{2}_{f_{i}} \log p(y_{i} \mid f(\mathbf{x}_{i})) \right] \mathbf{k}_{\mathbf{z}i}^{T}
% \end{align}

% weight space posterior in \cref{eq-laplace-approx-weight-space} in
% function space by linearising the BNN and interpreting it as a GP.
% \begin{align} \label{eq-laplace-approx-function-space}
%   \transitionFn_{\transitionParams_{i}}(\cdot) &\sim \mathcal{N} \left( \mu_{\transitionParams_{i}^{*}}(\cdot), K_{\transitionParams_{i}^{*}}(\cdot, \cdot') \right) \quad
%   \mu_{\transitionParams_{i}^{*}}(\cdot) &= \transitionFn_{\transitionParams^{*}_{i}}(\cdot) \quad
%   K_{\transitionParams_{i}^{*}}(\cdot, \cdot') &= \mathbf{J}_{\transitionParams^{*}_{i}}(\cdot) \bm\Sigma_{\transitionParams^{*}_{i}} \mathbf{J}_{\transitionParams^{*}_{i}}(\cdot')^{T}
% \end{align}
% where the Jacobian is given by $\mathbf{J}_{\transitionParams^{*}_{i}}(\cdot) = \odv{\transitionFnWithParams(\cdot)}{\transitionParams}_{\transitionParams=\transitionParams_{i}^{*}}$.

% \begin{align} \label{eq-laplace-approx-jacobian}
% \mathbf{J}_{\transitionParams^{*}_{i}}(\cdot) = \odv{\transitionFnWithParams(\cdot)}{\transitionParams}_{\transitionParams=\transitionParams_{i}^{*}}
% \end{align}

% % Note that the posterior in \cref{eq-laplace-approx-weight-space} is in weight space.
% We can represent the weight space posterior in \cref{eq-laplace-approx-weight-space} in
% function space by linearising the BNN and interpreting it as a GP.
% \begin{align} \label{eq-laplace-approx-function-space}
%   \transitionFn_{\transitionParams_{i}^{*}}(\cdot) &\sim \mathcal{N} \left( \mu_{\transitionParams_{i}^{*}}(\cdot), K_{\transitionParams_{i}^{*}}(\cdot, \cdot') \right) \quad
%   \mu_{\transitionParams_{i}^{*}}(\cdot) &= \transitionFn_{\transitionParams^{*}_{i}}(\cdot) \quad
%   K_{\transitionParams_{i}^{*}}(\cdot, \cdot') &= \mathbf{J}_{\transitionParams^{*}_{i}}(\cdot) \bm\Sigma_{\transitionParams^{*}_{i}} \mathbf{J}_{\transitionParams^{*}_{i}}(\cdot')^{T}
% \end{align}
% where the Jacobian is given by $\mathbf{J}_{\transitionParams^{*}_{i}}(\cdot) = \odv{\transitionFnWithParams(\cdot)}{\transitionParams}_{\transitionParams=\transitionParams_{i}^{*}}$.

% \begin{align} \label{eq-laplace-approx-jacobian}
% \mathbf{J}_{\transitionParams^{*}_{i}}(\cdot) = \odv{\transitionFnWithParams(\cdot)}{\transitionParams}_{\transitionParams=\transitionParams_{i}^{*}}
% \end{align}

\subsection{Fast Updates}
Importantly, \cite{changFantasizingDualGPs2022} show that in the dual space,
conditioning on new observations reduces to suming the dual variables from the previous time
step with an update, given by,
\todo{pick best way to show new data (using superscript new or just time indexing?)}
\begin{align} \label{eq-dual-update-svgp}
\dualParam{1}^{t+1} &\leftarrow \dualParam{1}^{t} +  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{t}, \action_{t}))} \left[ \log p(\state_{t+1} \mid \latentFn(\state_{t}, \action_{t}) ) \right] \\
\dualParam{2}^{t+1} &\leftarrow \dualParam{2}^{t} +  \nabla_{\meanParam{2}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{t}, \action_{t}))}  \left[ \log p(\state_{t+1} \mid \latentFn(\state_{t}, \action_{t}) ) \right]
\end{align}
Importantly, for a single new observation $((\state_{t}, \action_{t}), \state_{t+1})$ this update has
complexity $\mathcal{O}(\numInducing^{2})$.
This is a significant improvement to naive GP conditioning, which has complexity $\mathcal{O}((\numDataNew + \numDataOld)^{3})$
and sparse GP conditioning, which has complexity $\mathcal{O}((\numDataNew + \numDataOld)\numInducing^{2})$.
\todo{double check these complexities are right and cite them/show equaitons}
It is worth highlighting that the complexity of \cite{changFantasizingDualGPs2022} update does not increase during an episode.

\subsubsection{Model-based RL}

\subsubsection{Efficiently Sampling Functions for Posterior Sampling}
\cite{wilsonEfficiently2020}
\cite{wilsonPathwise2021}

\section{Model-based Reinforcement Learning with Fast Updates}

\subsection{Problem Statement and Background} \label{sec:problem-statement}
We consider environments with states \(\state \in \stateDomain \), actions \(\action \in \actionDomain\) and transition dynamics \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \) subject to
iid noise \(\noise_{t}\), given by,
\begin{align}
\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}.
\end{align}
\begin{assumption}
  (system properties) The environment's dynamics are $L_{\transitionFn}\text{-Lipschitz}$ continuous and the transition noise $\noise_{t}$ is $\sigma\text{-sub-Gaussian}$ for all $t \geq 0$.
\end{assumption}
